---
title: "p8105_hw3_sl4655"
author: "Shuya Liu"
date: "October 12, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggridges)
library(patchwork)
library(p8105.datasets)
```

## _Problem 1_

```{r}
data("instacart")
instacart <- instacart %>%
  janitor::clean_names()
```

#### A Short Description

The dataset has `r nrow(instacart)` observations of `r ncol(instacart)` variables: `r colnames(instacart)`. The descriptions for the variables are as below:

* order_id: order identifier
* product_id: product identifier
* add_to_cart_order: order in which each product was added to cart
* reordered: 1 if this prodcut has been ordered by this user in the past, 0 otherwise
* user_id: customer identifier
* eval_set: which evaluation set this order belongs in (Note that the data for use in this class is exclusively from the “train” eval_set)
* order_number: the order sequence number for this user (1=first, n=nth)
* order_dow: the day of the week on which the order was placed
* order_hour_of_day: the hour of the day on which the order was placed
* days_since_prior_order: days since the last order, capped at 30, NA if order_number=1
* product_name: name of the product
* aisle_id: aisle identifier
* department_id: department identifier
* aisle: the name of the aisle
* department: the name of the department

Take the first line as an example: the user with user ID 112108 placed an order with order ID 1. The first item he/she added to the card is Bulgarian Yogurt with product ID 49302, and this product has been ordered in the past. This is the 4th order of this user, and he/she placed this order on a Thursday at 10 a.m. It has been 9 days since his/her last order. The Bulgarian Yogurt is on the "yogurt" aisle with ID 120, in the department of dairy eggs with deparment ID 16. 

#### How many aisles are there, and which aisles are the most items ordered from?

```{r aisles}
# Number of aisles
n_aisles <- 
  instacart %>% 
  distinct(aisle_id) %>% 
  nrow()
```

* There are `r n_aisles` aisles.

* The top 6 aisles that the most items are ordered from are
```{r number_of_items}
# Count number of items sold on each aisle and arrange them in a descending order
instacart %>% 
  group_by(aisle) %>% 
  summarise(number_of_items = n()) %>%
  arrange(desc(number_of_items)) %>% 
  head() %>% 
  knitr::kable()
```

#### Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r aisle_plot}
instacart %>%
  group_by(aisle) %>%
  summarize(number_of_items = n()) %>%
  filter(number_of_items > 10000) %>%
  arrange(desc(number_of_items)) %>%
  ggplot(aes(x = aisle, y = number_of_items)) +
  geom_col() +
  geom_text(aes(label = number_of_items, angle = 90), size = 2, vjust = 0.2) +
  labs(
    title = "The Number of Items Ordered in Each Aisle",
    x = "Aisle",
    y = "Number of Items"
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 0.95, vjust = 0.2))
```

#### Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
```{r three_aisle}
instacart %>%
  filter(aisle == c('baking ingredients', 'dog food care', 'packaged vegetables fruits')) %>%
  group_by(aisle, product_name) %>% 
  summarise(num_of_purchase = n()) %>% 
  top_n(3) %>%
  knitr::kable()
```

#### Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
```{r}
mean_hour <- instacart %>% 
  filter(product_name == c('Pink Lady Apples', 'Coffee Ice Cream')) %>% 
  select(order_dow, order_hour_of_day, product_name) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_time = mean(order_hour_of_day)) %>% 
  pivot_wider(names_from = product_name, values_from = mean_time) %>% 
  t() %>%
  round(2) %>% 
  as.data.frame()
colnames(mean_hour) <- c('Sunday','Monday','Tuesday','Wednesday','Thursday','Friday','Saturday')
knitr::kable(mean_hour[-1,])
```

## _Problem 2_

#### Data Cleaning

* format the data to use appropriate variable names;
* focus on the “Overall Health” topic
* include only responses from “Excellent” to “Poor”
* organize responses as a factor taking levels ordered from “Poor” to “Excellent”

```{r data_cleaning}
data("brfss_smart2010")
df_brfss <- brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == 'Overall Health') %>%
  mutate(response = as.factor(response))
df_brfss$response <-
  factor(df_brfss$response,levels(df_brfss$response)[c(4,2,3,5,1)])
```

#### In 2002, which states were observed at 7 or more locations? What about in 2010?
```{r year_2002}
year_2002 <- df_brfss %>%
  filter(year == 2002) %>%
  group_by(locationabbr) %>%
  mutate(number_of_obs = length(unique(locationdesc))) %>%
  filter(number_of_obs == 7) %>% 
  distinct(locationabbr)
```

For year 2002, the states that were observed at 7 or more location were `r year_2002$locationabbr`.

```{r year_2010}
year_2010 <- df_brfss %>%
  filter(year == 2010) %>%
  group_by(locationabbr) %>%
  mutate(number_of_obs = length(unique(locationdesc))) %>%
  filter(number_of_obs == 7) %>% 
  distinct(locationabbr)
```

For year 2010, the states that were observed at 7 or more location were `r year_2010$locationabbr`.

#### Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state

```{r mean_data_value}
df_brfss %>%
  filter(response == 'Excellent') %>%
  group_by(year, locationabbr) %>%
  summarise(mean_data_value = mean(data_value)) %>% 
  mutate(state = locationabbr,
         mean_data_value = round(mean_data_value, 2)) %>% 
  select(year, state, mean_data_value) %>%
  ggplot(aes(x = year, y = mean_data_value, color = state)) +
  geom_line()
```

#### Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State

```{r two_panel}
year_2006 <- df_brfss %>%
  filter(year == 2006,
         locationabbr == 'NY') %>%
  ggplot(aes(x = response, y = data_value)) +
  geom_point() +
  ylim(0, 43) +
  labs(
    title = "Distribution of Data Value in NY in 2006",
    x = "Response",
    y = "Data Value"
  ) +
  theme(plot.title = element_text(size = 12))

year_2010 <- df_brfss %>%
  filter(year == 2010,
         locationabbr == 'NY') %>%
  ggplot(aes(x = response, y = data_value)) +
  geom_point() +
  ylim(0, 43) +
  labs(
    title = "Distribution of Data Value in NY in 2010",
    x = "Response",
    y = "Data Value"
  ) +
  theme(plot.title = element_text(size = 12))

year_2006 + year_2010
```

## _Problem 3_

#### Load, tidy, and wrangle the data

```{r load_data}
acc_data <- read_csv(file = "./Data/accel_data.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute",
    names_prefix = "activity_",
    values_to = "activity_per_minute") %>% 
  mutate(minute_id = as.numeric(minute),
         hour = minute_id %/% 60,
         minute = minute_id %% 60,
         day = day_id %% 7,
         week = as.factor(week),
         day = factor(day, 
                      levels = c(0:6), 
                      labels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))) %>%
  select(week, day, hour, minute, activity_per_minute, minute_id)
```

This dataset has `r nrow(acc_data)` observations of `r ncol(acc_data)` variables:`r colnames(acc_data)`.

#### Total Activity over the Day

```{r total_activity_per_day}
acc_data %>%
  group_by(week, day) %>% 
  summarise(total_activity = round(sum(activity_per_minute))) %>%
  pivot_wider(names_from = day, values_from = total_activity) %>%
  knitr::kable(caption = "Total Activity per Day")
```

Basically, the patient has a relatively constant amount of activities per day, expect for some days that he moved too much or too little. On the Tuesday of the first week, the patient barely moved. On the Wednesdays of the last 2 weeks, the patient barely moved.

#### Make a single-panel plot that shows the 24-hour activity time courses for each day
```{r activity_plot}
acc_data %>%
  group_by(week, day) %>% 
  ggplot(aes(x = minute_id, y = activity_per_minute, color = day)) +
  geom_line(alpha = .5) + 
  xlim(0, 1440) +
  facet_grid(week ~ .) +
  labs(
    title = "The Activity over the Course of the Day",
    x = "Time of the Day",
    y = "Amount of Activities"
  )
```

As we can see the graph, the patient's total amount of activties decreases as the week passes. For Mondays, the patient has more activites around 20:00 to 21:30. For Thursdays, the patients has more activities around 10:30 to 12:30. The activities tend to decrease to its minimum after 23:00, and starts to increase after 5:00.